Target Markdown

### 2023-10-11 _targets.Rmd
Compute stability-richness slope vs other variables

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("tidyverse", "magrittr", "cowplot", "here", "arrow", "piecewiseSEM", "semEff", "easystats", "glmmTMB", "ggcorrplot", "kableExtra"))
source(here::here("R", "network_metrics.R"))
source(here::here("R", "string_replacements.R"))
source(here::here("R", "stability_decomposition.R"))
source(here::here("R", "model_helpers.R"))
source(here::here("R", "sem_helpers.R"))
```

@@

@@

```{r, message = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
library(ggcorrplot)
source(here::here("R", "network_metrics.R"))
source(here::here("R", "string_replacements.R"))
source(here::here("R", "stability_decomposition.R"))
source(here::here("R", "model_helpers.R"))
```

@@

```{r}
sim_fw %>%
  filter(
    rho %in% c(0), productivity == 5, env_stoch == .5
    ) %>%
  ggplot(aes(y = stab_pop, x = as.factor(Z))) +
  geom_boxplot()
```

```{r}
sim_fw %>%
  filter(
    rho %in% c(0), productivity == 5, env_stoch == .5,
    Z == 500
    ) %>%
  pivot_longer(c(stab_com, async, sae_total, cpe,
      sae_even, evenness_sae,
      cpe_int, cpe_env, stab_pop
      ), names_to = "async_decomp", values_to = "values") %>%
  mutate(async_decomp = var_replacement()[async_decomp]) %>%
  ggplot(aes(x = as.factor(richness), y = values)) +
  geom_boxplot() +
  labs(x = "Connectance", y = "Asynchrony component") +
  facet_wrap(~async_decomp, ncol = 3, scales = "free_y") +
  theme_bw()
```

@@

sim_sp %>%
  filter(env_stoch == .3, rho == 1, Z == 100) %>%
  ggplot(aes(x = tlvl, y = log(cv_sp))) +
  geom_point() +
  geom_smooth()
```

@@

```{r}
p_pred_sr_avg_tlvl_h <- pred_data %>%
  filter(
    ct_alive %in% c(.1),
    h == c(3),
    rho %in% c(0, .75),
    Z == 10,
    avg_int_strength %in% c(.1),
    w_avg_tlvl == c(1.5, 3)) %>%
  ggplot(aes(x = richness, y = stab_com,
      linetype = as.factor(w_avg_tlvl), color = as.factor(rho))) +
  geom_line() +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  theme_bw()
ggsave(
  here("report", "figures", "p_pred_sr_avg_tlvl_h.png"),
  plot = p_pred_sr_avg_tlvl,
  scale = 2.4,
  width = 88,
  height = 88 * 1,
  units = "mm"
)
```

@@

```{r}
tar_load(sim_fw_sem)
sim_fw_sem %<>% mutate(richness = sqrt(richness))
ti <- as.psem(sem_model_list)
update(ti, avg_omnivory %~~% w_avg_tlvl)
sem_coeff <- summary(ti)$coefficients[, c(1, 2, 8, 9)]
map_dfr(sem_model_list, rsquared) %>%
      mutate(Response = var_replacement()[Response]) %>%
      select(Response, R.squared)
```

```{r}
sem_model_list <- list(
      ct_alive = lm(ct_alive ~ ct + S, sim_fw_sem),
      richness = lm(richness ~ ct + S, sim_fw_sem),
      #max_tlvl = lm(max_tlvl ~ ct + S, sim_fw_sem),
      #w_avg_tlvl = lm(w_avg_tlvl ~ ct + S + Z, sim_fw_sem),
      #avg_omnivory = lm(avg_omnivory ~ ct + S, sim_fw_sem),
      avg_int_strength = lm(avg_int_strength ~ ct + S + h + Z, sim_fw_sem),
      pop_stab = lm(pop_stab ~ richness + avg_int_strength + ct_alive + env_stoch, sim_fw_sem),
      evenness_sae = lm(evenness_sae ~ richness + avg_int_strength + ct_alive, sim_fw_sem),
      sae_even = lm(sae_even ~ richness, sim_fw_sem),
      cpe_int = lm(cpe_int ~ richness + avg_int_strength + ct_alive, sim_fw_sem),
      cpe_env = lm(cpe_env ~ richness + rho, sim_fw_sem),
      async = lm(async ~ sae_even + evenness_sae + cpe_env + cpe_int, sim_fw_sem),
      stab_com = lm(stab_com ~ pop_stab + async, sim_fw_sem)
    )
ti <- as.psem(sem_model_list)
update(ti, avg_omnivory %~~% w_avg_tlvl)
sem_coeff <- summary(ti)$coefficients[, c(1, 2, 8, 9)]
map_dfr(sem_model_list, rsquared) %>%
      mutate(Response = var_replacement()[Response]) %>%
      select(Response, R.squared)
```

>>>>

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("tidyverse", "magrittr", "cowplot", "here", "arrow", "piecewiseSEM", "semEff", "easystats", "glmmTMB", "ggcorrplot", "kableExtra"))
# Load custom functions
lapply(list.files(here::here("R"), full.names = TRUE), source)
```

@@

```{targets simplified-sem}
list(
  tar_target(sem_simplified_model_list,
    list(
      ct_alive = lm(ct_alive ~ ct + S, sim_fw_sem),
      richness = lm(richness ~ ct + S, sim_fw_sem),
      #max_tlvl = lm(max_tlvl ~ ct + S + Z, sim_fw_sem),
      w_avg_tlvl = lm(w_avg_tlvl ~ ct + S + Z, sim_fw_sem),
      avg_omnivory = lm(avg_omnivory ~ ct + S, sim_fw_sem),
      avg_int_strength = lm(avg_int_strength ~ ct + S + h + Z, sim_fw_sem),
      pop_stab = lm(pop_stab ~ richness + avg_int_strength + w_avg_tlvl + ct_alive + env_stoch + h + Z, sim_fw_sem),
      sae_total = lm(sae_total ~ richness + avg_int_strength + w_avg_tlvl + ct_alive + h + Z, sim_fw_sem),
      cpe = lm(cpe ~ richness + avg_int_strength + w_avg_tlvl + ct_alive + h + Z + rho, sim_fw_sem),
      async = lm(async ~ sae_total + cpe, sim_fw_sem),
      stab_com = lm(stab_com ~ pop_stab + async, sim_fw_sem)
    )
    ),
  tar_target(sem_simplified, {
    ti <- as.psem(sem_simplified_model_list)
    update(ti, avg_omnivory %~~% w_avg_tlvl)
    }
    ),
  tar_target(sem_simplified_coeff, {
    ti <- summary(sem_simplified)$coefficients[, c(1, 2, 8, 9)]
    colnames(ti)[4] <- "Signif"
    ti
    }),
  tar_target(sem_simplified_collinearity,
    map_dfr(setNames(sem_simplified_model_list, names(sem_simplified_model_list)),
      performance::check_collinearity,
      .id = "response") %>%
    as_tibble),
  tar_target(sem_simplified_rsquared,
    map_dfr(sem_simplified_model_list, rsquared) %>%
      mutate(Response = var_replacement()[Response]) %>%
      select(Response, R.squared)
    ),
  tar_target(tab_sem_simplified_vif,
    sem_simplified_collinearity %>%
      mutate(
        response = var_replacement()[response],
        Term = var_replacement()[Term]
        ) %>%
      kable(
        caption = "Variance Inflation Factors for the Structural Equation Model.") %>%
      kable_styling()
    ),
  tar_target(tab_sem_simplified_rsquared,
    sem_simplified_rsquared %>%
      kable(caption = "Rsquared for the Structural Equation Model.") %>%
      kable_styling()
  )
)
```

```{targets model-to-pred}
list(
  tar_target(model_fw_stab,
    glmmTMB(formula_stab_fw_model(
        rand_effect = TRUE,
        term_to_del = "richness : rho : Z +"),
      data = mutate(sim_fw, across(c(stab_com, richness), log))
      )),
  tar_target(model_fw_stab_scaled,
    glmmTMB(formula_stab_fw_model(
        rand_effect = TRUE,
        term_to_del = "richness : rho : Z +"
        ),
      data = sim_fw %>%
        select(-productivity) %>%
        mutate(across(c(stab_com, richness), log)) %>%
        mutate(across(c(sim_id, fw_id), as.factor)) %>%
        mutate(
          across(c(stab_com, richness,
              env_stoch, rho, Z,
              ct_alive, w_avg_tlvl, avg_int_strength, avg_omnivory),
            ~scale(.x)[, 1])
        )
        )
    ),
  tar_target(model_fw_stab_scaled_coeff,
    confint(tar_read(model_fw_stab_scaled), level = 0.95) %>%
      as.data.frame() %>%
      rownames_to_column("term") %>%
      as_tibble() %>%
      mutate(term = str_replace(term, "cond\\.", ""))
    ),
  tar_target(pred_data, {
    to <- list(
      richness = log(seq(1, 30, 1)),
      ct_alive = c(.05, .10, .30),
      rho = c(0, .75, 1),
      avg_int_strength = c(.01, .02, .05, .1),
      env_stoch = c(.1, .6),
      w_avg_tlvl = c(1.5, 2, 3),
      Z = c(10, 100),
      h = c(2, 3),
      fw_id = NA) %>% expand.grid()
    ti <- predict(model_fw_stab,
      se.fit = TRUE, re.form = NA,
      newdata = to)
    to$stab_com <- ti$fit
    to$stab_com_se <- ti$se.fit
    to$resp_div <- rho_replacement()[as.character(to$rho)]
    to
    }),
  tar_target(fix_fw_stab, fixef(model_fw_stab)$cond),
  tar_target(pred_sr_rho_int,
    list(
      rho = seq(0, 1, .02),
      avg_int_strength = seq(0.01, 0.2, .01)
      ) %>%
    expand.grid() %>%
    mutate(
      slope_richness = map2_dbl(rho, avg_int_strength,
        ~get_slope_change(
          coeff = fix_fw_stab,
          gen_term = "richness",
          coeff_values = c(
            "rho" = .x,
            "env_stoch" = .1,
            "ct_alive" = .2,
            "avg_int_strength" = .y,
            "h" = 2,
            "w_avg_tlvl" = 1.5,
            "Z" = 100
          )
        )
      )
    )
  )
)
```

@@

```{r, message = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
library(ggcorrplot)
lapply(list.files(here::here("R"), full.names = TRUE), source)
```

```{r}

# plot just the interactions
# TODO: order by values
ti %>%
  filter(str_detect(term, ":")) %>%
  ggplot(aes(x = Estimate, y = term)) +
  geom_point()
```
```{r}
p_pred_sr_rho_int <- tar_read(pred_sr_rho_int) %>%
  ggplot(aes(x = rho, y = avg_int_strength)) +
  geom_raster(aes(fill = slope_richness)) +
  scale_fill_gradient2(
    low = "red",
    mid = "white",
    high = "blue",
    midpoint = 0,
    space = "Lab",
    na.value = "grey50",
    guide = "colourbar",
    aesthetics = "fill") +
  labs(
    x = "Environmental correlation",
    y = "Avg interaction strength",
    fill = "Slope Stability-Richness"
  ) +
theme(legend.position = "bottom")
ggsave(
  here("paper", "figures", "p_pred_sr_rho_int.pdf"),
  plot = p_pred_sr_rho_int,
  scale = 2.4,
  width = 88,
  height = 88 * 1,
  units = "mm"
)

```{r}
tar_load(pred_data)
pred_data %>%
  filter(
    ct_alive %in% c(.1),
    rho %in% c(0, .75),
    richness >= 1,
    Z == 10,
    h %in% c(3),
    avg_int_strength %in% c(.01, .1),
    w_avg_tlvl == 1.5
    ) %>%
  ggplot(aes(
      x = richness, y = stab_com,
      color = as.factor(resp_div),
      linetype = as.factor(avg_int_strength),
      alpha = as.factor(avg_int_strength)
      )) +
  ylim(2, 8) +
  geom_line(linewidth = 1) +
  scale_alpha_manual(values=c(1, 1), guide = "none") +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  labs(
    x = "Species richness (Log)", y = "Community stability (Log)",
    linetype = "Avg. Interaction strength",
    color = "Response diversity"
  ) +
  theme_bw() +
  theme(legend.position = "bottom")

#pred_data$async <- predict(mod_tot_async, newdata = pred_data)
#pred_data$pop_stab <- predict(mod_tot_pop_stab, newdata = pred_data)
```

## Simplified SEM

simplified SEM for main text

```{r}
tar_read(sem_simplified_coeff)
```

```{r}
tar_load(sim_fw)
colnames(sim_fw)
length(unique(sim_fw$fw_id))
```

## Modelling interaction between stability and food-web structure

```{r}
model_fw_stab <- glmmTMB(formula_stab_fw_model(
    rand_effect = TRUE,
    term_to_del = "richness : rho : Z +"),
  data = mutate(sim_fw, across(c(stab_com, richness), log))
)
diagnose(model_fw_stab)

model_fw_stab_scaled <- glmmTMB(formula_stab_fw_model(rand_effect = TRUE),
  data = sim_fw %>%
    select(-productivity) %>%
    mutate(across(c(stab_com, richness), log)) %>%
    mutate(across(c(sim_id, fw_id), as.factor)) %>%
    mutate(
      across(c(stab_com, richness,
          env_stoch, rho, Z,
          ct_alive, w_avg_tlvl, avg_int_strength, avg_omnivory),
        ~scale(.x)[,1])
    )
)
fixef(model_fw_stab_scaled)
diagnose(model_fw_stab_scaled)
```

```{r}
mod_tot <- lm(stab_com ~
  richness + rho + env_stoch +
  ct_alive + w_avg_tlvl + Z +
  avg_int_strength +
  h +
  richness * rho +
  richness * env_stoch +
  richness * ct_alive +
  richness * avg_int_strength +
  richness * h +
  richness * w_avg_tlvl +
  richness * Z +
  richness * rho * ct_alive +
  richness * rho * w_avg_tlvl +
  richness * rho * avg_int_strength +
  richness * rho * h +
  richness * rho * Z +
  richness * env_stoch * ct_alive +
  richness * env_stoch * Z +
  richness * env_stoch * avg_int_strength +
  richness * env_stoch * h +
  richness * env_stoch * w_avg_tlvl,
  mutate(sim_fw, across(c(stab_com, richness), log)) #, richness
  )
```

##

@@

```{r}
sim_fw %>%
  filter(
    rho %in% c(0), env_stoch == .3
    ) %>%
  ggplot(aes(y = stab_pop, x = as.factor(Z))) +
  geom_boxplot()
```

```{r}
sim_fw %>%
  filter(
    rho %in% c(0), env_stoch == .3, Z == 25) %>%
  pivot_longer(c(stab_com, async, sae_total, cpe,
      sae_even, evenness_sae,
      cpe_int, cpe_env, stab_pop
      ), names_to = "async_decomp", values_to = "values") %>%
  mutate(async_decomp = var_replacement()[async_decomp]) %>%
  ggplot(aes(x = as.factor(richness), y = values)) +
  geom_boxplot() +
  labs(x = "Species richness", y = "Asynchrony component") +
  facet_wrap(~async_decomp, ncol = 3, scales = "free_y") +
  theme_bw()
```

@@

sim_sp %>%
  filter(env_stoch == .4, rho == 1, Z == 100) %>%
  ggplot(aes(x = tlvl, y = log(cv_sp))) +
  geom_point() +
  geom_smooth()
```

@@

```{r}
p_pred_sr_avg_tlvl_h <- pred_data %>%
  filter(
    ct_alive %in% c(.1),
    h == c(3),
    rho %in% c(0, .75),
    Z == 10,
    avg_int_strength %in% c(.1),
    w_avg_tlvl == c(1.5, 3)) %>%
  ggplot(aes(x = richness, y = stab_com,
      linetype = as.factor(w_avg_tlvl), color = as.factor(rho))) +
  geom_line() +
  facet_grid(cols = vars(env_stoch), labeller = label_both) +
  theme_bw()
ggsave(
  here("report", "figures", "p_pred_sr_avg_tlvl_h.png"),
  plot = p_pred_sr_avg_tlvl_h,
  scale = 2.4,
  width = 88,
  height = 88 * 1,
  units = "mm"
)
```

@@
